---
title: "AI Measurement Paper — Full Psychometric Pipeline"
output: html_notebook
---

This notebook runs the complete E1–E5 psychometric pipeline (plus the E1b
model-ability extension) end to end, by sourcing the actual production
scripts under `code/` rather than duplicating their logic here. There is a
single source of truth: `code/*.R`. If those scripts change, re-running this
notebook picks up the change automatically.

Default data stream is the real Delphi stream (`results/r_data`), matching
`code/run_psychometrics.R`. A final optional section reruns E3–E5 against the
automated-score stream (`results/r_data_dev`) for comparison.

## Setup

```{r setup}
suppressPackageStartupMessages({
  library(knitr)
})

# Robustly locate the repo root regardless of how/where this notebook is run
# from (knit button, Run All, rmarkdown::render() from an arbitrary working
# directory, etc.) -- do NOT rely on commandArgs()/--file= (there is none
# when a script is source()'d) or on knitr's working-directory default
# (inconsistent across invocation methods). Instead walk up from both the
# document's own location and the current working directory looking for the
# sibling `code/` and `results/` directories that mark the repo root.
find_repo_root <- function() {
  candidates <- character(0)
  doc_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) NULL)
  if (!is.null(doc_path) && nzchar(doc_path)) candidates <- c(candidates, dirname(doc_path))
  candidates <- c(candidates, getwd())

  for (start in candidates) {
    d <- normalizePath(start, mustWork = FALSE)
    for (i in 1:6) {
      if (dir.exists(file.path(d, "code")) && dir.exists(file.path(d, "results"))) return(d)
      parent <- normalizePath(file.path(d, ".."), mustWork = FALSE)
      if (identical(parent, d)) break
      d <- parent
    }
  }
  stop("Could not locate the AI_measurement_paper repo root ",
       "(looked for sibling code/ and results/ directories).")
}

ROOT <- find_repo_root()
RES    <- file.path(ROOT, "results")
FIGS   <- file.path(RES, "figures")
R_DATA <- file.path(RES, "r_data")
cat("Repo root:", ROOT, "\n")

# Each code/e*.R script computes its own ROOT via a --file=/commandArgs()
# trick that only works when invoked as `Rscript code/foo.R`. Sourcing a
# script from a notebook chunk never sets --file=, so those scripts fall
# back to `normalizePath("..")` -- which resolves relative to the current
# working directory at that moment. run_script() makes that fallback
# resolve correctly (to ROOT) by temporarily setting the working directory
# to code/ for the duration of the source() call, exactly matching what
# `Rscript code/foo.R` run from the repo root would see. This avoids having
# to modify the production scripts at all.
run_script <- function(filename, env_vars = list()) {
  old_wd  <- getwd()
  old_env <- Sys.getenv(names(env_vars), unset = NA)
  on.exit({
    setwd(old_wd)
    for (i in seq_along(env_vars)) {
      nm <- names(env_vars)[i]
      if (is.na(old_env[nm])) Sys.unsetenv(nm) else Sys.setenv(!!nm := old_env[nm])
    }
  }, add = TRUE)

  if (length(env_vars) > 0) do.call(Sys.setenv, env_vars)
  setwd(file.path(ROOT, "code"))
  source(filename, echo = FALSE)
}

show_csv <- function(name) {
  path <- file.path(RES, name)
  if (!file.exists(path)) { cat("(not found:", name, ")\n"); return(invisible(NULL)) }
  knitr::kable(read.csv(path, stringsAsFactors = FALSE))
}
```

## E1 — Factor structure (automated + Delphi criteria)

```{r e1}
run_script("e1_factor_structure.R")
```

```{r e1-results}
show_csv("T1_cfa_fit_indices.csv")
show_csv("T1_mirt_bic.csv")
show_csv("T1_factor_loadings.csv")
```

## E1b — IRT-weighted model ability (theta), reusing E1's MIRT fit

```{r e1b}
run_script("e1b_model_theta.R")
```

```{r e1b-results}
show_csv("T1b_model_theta.csv")
show_csv("T1b_theta_vs_naive.csv")
```

## E2 — Variance decomposition (MFRM)

```{r e2}
run_script("e2_variance_decomp.R")
```

```{r e2-results}
show_csv("T2_variance_components_fixed_criterion.csv")
show_csv("T2_variance_components.csv")
show_csv("T2_variance_components_zscored.csv")
show_csv("T2_reliability.csv")
show_csv("T2_convergence_by_round.csv")
```

## E3 — Model stringency and prompt generatability (flipped 2PL IRT)

```{r e3}
run_script("e3_item_params.R")
```

```{r e3-results}
show_csv("T3_model_parameters.csv")
```

## E4 — Objaverse training-data sensitivity

```{r e4}
run_script("e4_dif_analysis.R")
```

```{r e4-results}
sens <- read.csv(file.path(RES, "T4_objaverse_sensitivity.csv"))
cat("N items:", nrow(sens), "\n")
cat("Mean |gap|:", round(mean(sens$abs_gap, na.rm = TRUE), 3), "\n")
cat("% items with |gap| >= 0.40:", round(100 * mean(sens$abs_gap >= 0.40, na.rm = TRUE), 1), "\n")
cat("Complete reversals (|gap| == 1.0):", sum(sens$abs_gap == 1.0, na.rm = TRUE), "\n")
```

## E5 — Adaptive 30-item subset (tinyEval3D)

```{r e5}
run_script("e5_adaptive_subset.R")
```

```{r e5-results}
show_csv("T5_subset_comparison.csv")
```

## Optional: rerun E3–E5 against the automated-score stream for comparison

Uses the `E{3,4,5}_R_DATA_DIR` / `E{3,4,5}_OUT_SUFFIX` overrides so the
default Delphi-stream results above are untouched; outputs are written with
an `_automated` suffix.

```{r automated-stream, eval=dir.exists(file.path(RES, "r_data_dev"))}
run_script("e3_item_params.R", list(E3_R_DATA_DIR = "r_data_dev", E3_OUT_SUFFIX = "_automated"))
run_script("e4_dif_analysis.R", list(E4_R_DATA_DIR = "r_data_dev", E4_OUT_SUFFIX = "_automated"))
run_script("e5_adaptive_subset.R", list(E5_R_DATA_DIR = "r_data_dev", E5_OUT_SUFFIX = "_automated"))
```

```{r automated-stream-results, eval=dir.exists(file.path(RES, "r_data_dev"))}
show_csv("T3_model_parameters_automated.csv")
show_csv("T5_subset_comparison_automated.csv")
```

## Session info

```{r session-info}
sessionInfo()
```
